A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-03-24 CC/writeup.Rnw
add pdf

%\definecolor{Blue}{rgb}{0,0,0.5}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\sct}[1]{{\color{purple} #1}}
\newcommand{\sctq}[1]{{\color{red} #1}} %things in red are questions for you or concerns that i have.

@@

\maketitle
\begin{abstract}
Abstract will be here
\end{abstract}
<<chunk_setup, echo=FALSE>>=
library(knitr)
opts_chunk$set(tidy = FALSE, echo=FALSE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
@
<<readingdata, eval=TRUE>>=
 library(ggplot2) # needs to be the version with geom_net!
 theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
 data(blood, football, theme_elements, lesmis)
 load("data/madmen.rda") # relative path should work for both of us
names(madmen$vertices)[1] <- "label"
names(blood$vertices)[1] <- "label"
names(theme_elements$vertices)[1] <- "label"
football$vertices$label <- as.factor(football$vertices$label)
football$vertices$value <- as.factor(football$vertices$value)
 email.net <- read.csv("data/email-data.csv")
 email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romatic relationship.}
\end{figure}

@@

\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
  \subsection{Madmen Network (again?? or no?)}
  \sctq{leaving blank for now.}
  \subsection{College Football}
  \emph{Note: would be best with coloring vertices by conference}
  This next example comes from M.E.J. Newman's network data web page \citep{football}.  It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.  There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = I(2), vlabel = TRUE,aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(value)])) +
  expand_limits(x = c(0,1), y = c(0,1)) +
  theme_net
@
\sct{
The network of football games is given in figure \ref{fig.cap:football_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
\end{figure}
  \subsection{Les Mis\'{e}rables}
  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
  This next network comes from \citet{lesmis}. It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.  There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.  There are 254 edges in this network. The edges are also weighted by the number of coappearances. The largest weighting is 31, between the characters Jean Valjean and Cosette.}
<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
\end{figure}
\sct{
  \subsection{Protein Interaction Network in Yeast}
  This example of a protein interaction network comes from \citet{protein}. It is the complete protein-protein interaction network in the yeast species \emph{S. cerevisiae}. There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.  These edges represent ``direct physical interactions" between any 2 proteins \citep[][p. 42]{protein}. 
}
<<read_data>>=
yeast <- read.table("data/protein.txt")
yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
@
<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
@

>>>>

%\definecolor{Blue}{rgb}{0,0,0.5}
\usepackage{color}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\sct}[1]{{\color{purple} #1}}
\newcommand{\sctq}[1]{{\color{red} #1}} %things in red are questions for you or concerns that i have.

@@

\maketitle
\begin{abstract}
Abstract will be here
\end{abstract}
<<chunk_setup, echo=FALSE>>=
library(knitr)
opts_chunk$set(tidy = FALSE, echo=FALSE, cache=TRUE, eval = TRUE, message = FALSE, warning = FALSE,  fig.height = 8, fig.width = 8, fig.align = 'center',fig.show='hide')
@
<<readingdata, eval=TRUE, echo=FALSE>>=
 library(ggplot2) # needs to be the version with geom_net!
 theme_net <- theme(aspect.ratio = 1, line = element_blank(), panel.background = element_blank(), plot.background = element_blank(), axis.text=element_blank(),axis.title=element_blank())
 data(blood, football, theme_elements, lesmis)
 load("data/madmen.rda") # relative path should work for both of us
names(madmen$vertices)[1] <- "label"
names(blood$vertices)[1] <- "label"
names(theme_elements$vertices)[1] <- "label"
football$vertices$label <- as.factor(football$vertices$label)
football$vertices$value <- as.factor(football$vertices$value)
 email.net <- read.csv("data/email-data.csv")
 email.names <- data.frame(label = as.factor(unique(c(as.character(email.net$From),as.character(email.net$to)))))

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/madmen_ex-1.pdf}
\caption{\label{fig.cap:madmen_ex} Graph of the characters in the show Mad Men who are linked by a romantic relationship.}
\end{figure}

@@

\sct{The inheritance structure is plotted in figure \ref{fig.cap:theme_ex}.
  \subsection{Madmen Network (again?? or no?)}
  \sctq{leaving blank for now.}
  \subsection{College Football}
  \emph{Note: would be best with coloring vertices by conference}
  This next example comes from M.E.J. Newman's network data web page \citep{football}.  It is an undirected network consisting of all regular season college football games played between Division I schools in Fall of 2000.  There are 115 vertices and 613 edges: each vertex represents a school, and an edge represents a game played between two schools.}
<<football_ex, echo=TRUE, fig.cap="All regular season Division I college football games in the Fall 2000 season.">>=
ggplot(data = football$edges, aes(from_id = from, to_id = to)) +
  geom_net(vertices = football$vertices, ealpha = 0.2, vsize = 5, vlabel = TRUE, aes(vcolour=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928")[as.numeric(factor(value))])) +
  expand_limits(x = c(0,1), y = c(0,1)) +
  theme_net
@
\sct{
The network of football games is given in figure \ref{fig.cap:football_ex}.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/football_ex-1.pdf}
\caption{\label{fig.cap:football_ex} The network of regular season Division I college football games in the season of fall 2000. The vertices and their labels are colored by conference.}
\end{figure}
  \subsection{Les Mis\'{e}rables}
  \emph{Note: data set has edge weightings for number of co-occurences of characters. esize would be useful here!} \\
  This next network comes from \citet{lesmis}. It is an undirected network of coappearances of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.  There are 77 vertices representing each of the 77 characters in the book, and an edge connects two vertices if those two characters appear in the same chapter of the book.  There are 254 edges in this network. The edges are also weighted by the number of coappearances. The largest weighting is 31, between the characters Jean Valjean and Cosette.}
<<lesmis_ex, echo=TRUE, fig.cap="Coappearance network of characters in Victor Hugo's \\emph{Les Mis\\'{e}rables}">>=
ggplot(data = lesmis$edges, aes(from_id = from, to_id = to)) + geom_net(vertices = lesmis$vertices, ealpha = 2/10, layout = 'fruchtermanreingold', vlabel=TRUE) + expand_limits(x = c(0,1.1), y = c(0,1)) + theme_net
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/lesmis_ex-1.pdf}
\caption{\label{fig.cap:lesmis_ex} Coappearance network of characters in Victor Hugo's \emph{Les Mis\'{e}rables}.}
\end{figure}
\sct{
  \subsection{Protein Interaction Network in Yeast}
  This example of a protein interaction network comes from \citet{protein}. It is the complete protein-protein interaction network in the yeast species \emph{S. cerevisiae}. There are 1,870 proteins that make up the vertices of this network, and there are 2,240 edges between them.  These edges represent ``direct physical interactions" between any 2 proteins \citep[][p. 42]{protein}.
}
<<read_data>>=
yeast <- read.table("data/protein.txt")
yeast <- list(edges = yeast, vertices = data.frame(label = min(yeast):max(yeast)))
@
<<yeast_ex, echo=TRUE, fig.cap="Protein in protein interation network in yeast.", cache=TRUE>>=
ggplot(data = yeast$edges, aes(from_id = V1, to_id = V2)) +
  geom_net(vertices = yeast$vertices, ealpha = 0.1, valpha = .5) +
  expand_limits(x = c(0,1), y = c(0,1)) + theme_net
@
